Augmenting the DonorsChoose.org Corpus for Meta-Learning

### 2019-04-08 doc/poster/main.tex
poster/main.tex: Redesign layout and shorten body
Shorten the text in the body of the poster, change the look
significantly and adapt countless smaller things to make the poster look
proper.

Prune dropped images and add the logo of the ADAPT group to the "res/"
folder instead.

\usepackage[inline]{enumitem}  % Inline enumeration capability
\usepackage[font=small,labelfont=bf]{caption}  % Customize caption aesthetics
\usepackage[font=small]{subcaption}  % Customize sub-captions

@@

@@

% Prettify poster
\definecolor{BlockTitle}{RGB}{32, 113, 182}
\definecolor{BlockBg}{RGB}{230, 230, 230}
\usefonttheme[onlymath]{serif}%
\setbeamercolor*{title}{fg = black}%
\setbeamercolor*{title page}{fg = black}%
\setbeamercolor*{block title}{fg = BlockTitle, bg = BlockBg}%
\setbeamercolor*{caption name}{fg = black}%
\setbeamercolor*{bibliography entry author}{fg = black}%
\setbeamercolor*{bibliography entry title}{fg = black}%
\setbeamercolor*{bibliography entry location}{fg = black}%
\setbeamercolor*{bibliography entry note}{fg = black}%
\setbeamerfont{title}{series = \bfseries}%
\setbeamerfont{institute}{size = \small}%
\setbeamertemplate{caption}[numbered]%
\setbeamertemplate{caption label separator}[endash]%
\setbeamertemplate{bibliography item}[text]
%\usebackgroundtemplate{\includegraphics[width=\paperwidth,height=\paperheight]{./MUW_green.pdf}}  % Background image

\title{Augmenting the DonorsChoose.org Corpus for Meta-Learning}
\subject{Algorithm Selection and Meta-Learning in Information Retrieval}
\author{Gordian~Edenhofer\inst{1,2} \and Andrew~Collins\inst{1} \and Akiko~Aizawa\inst{2} \and Joeran~Beel\inst{1,2}}
\institute{%
\inst{1}Trinity College Dublin, School of Computer Science \& Statistics, ADAPT Centre, Ireland\\
\inst{2}National Institute of Informatics Tokyo, Digital Content and Media Sciences Division, Japan
\and \href{mailto:beelj@tcd.ie}{beelj@tcd.ie}, \href{mailto:collinsa@tcd.ie}{collinsa@tcd.ie}, \href{mailto:aizawa@nii.ac.jp}{aizawa@nii.ac.jp}
}
\date{}  % Set date to nothing

@@

\begin{frame}[t,fragile=singleslide]{}
	% Headline
	\begin{columns}[t]
		\begin{column}{0.02\textwidth}
		\end{column}

		\begin{column}{0.18\textwidth}
			\flushleft%
			\vspace{1em}
			\includegraphics[width=0.8\columnwidth]{{{../res/Logo of Trinity College Dublin}}}
		\end{column}

		\begin{column}{0.6\textwidth}
			\vspace{1em}
			\titlepage%
		\end{column}

		\begin{column}{0.18\textwidth}
			\flushright%
			\vspace{1em}
			\vspace{2em}\includegraphics[width=0.8\columnwidth]{{{../res/Logo of NII}}}
		\end{column}

		\begin{column}{0.02\textwidth}
		\end{column}
	\end{columns}

	\begin{columns}[t]
		\begin{column}{0.45\textwidth}
			\begin{block}{Abstract}
				The DonorsChoose.org dataset of past donations provides a big and feature-rich corpus of users and items. The dataset matches donors to projects in which they might be interested in and hence is intrinsically about recommendations. Due to the availability of detailed item-, user- and transaction-features, this corpus represents a suitable candidate for meta-learning approaches to be tested. This study aims at providing an augmented corpus for further recommender systems studies to test and evaluate meta-learning approaches. In the augmentation, metadata of collaborative and content-based filtering techniques is amended to the corpus. It is further extended with aggregated statistics of users and transactions and an exemplary meta-learning experiment. The performance in the learning subsystem is measured via the recall of recommended items in a Top-N test set. The augmented dataset and the source code are released into the public domain at \href{https://github.com/BeelGroup/Augmented-DonorsChoose.org-Dataset}{GitHub:BeelGroup/Augmented-DonorsChoose.org-Dataset}.
			\end{block}

			\begin{block}{Introduction}
				Meta-Learning is the process of applying machine learning on metadata of previous machine learning experiments. The goal is to improve existing approaches by combining the strengths of several single machine learnings~\cite{AIR:MetalearningSurveyOfTrendsAndTechnologiesLemke2015}. The metadata though needs to be computed in advance.

				Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, the choice of the metric for computing metadata of single machine learners is important. Using the RMSE, as is done in some of the aforementioned publications, disallows using content-based methods.

@@

		\begin{column}{0.45\textwidth}
			\begin{block}{Methodology}
				The DonorsChoose.org dataset is composed of past donations featuring $4.69$ million transactions performed by $2.12$ million users having donated to $1.11$ million projects. The corpus provides significantly more transactions with more detailed item and transaction information compared to, e.g., the popular MovieLens dataset. An exemplary itemized view of the augmented dataset is shown in~\autoref{tab:IllustrativeTable} with the augmented data to the right of the separator.

				\vspace*{-.5\baselineskip}  % TODO: Fix this HACK
				\begin{table}[ht]
					\centering
					\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{{{../res/table}}}
					\caption{Illustrative example of the overall design of the augmented transaction table with amended learning subsystem performance scores, statistics and meta-learner information.}\label{tab:IllustrativeTable}
				\end{table}

				SVD and KNN are used as representatives for collaborative filtering techniques and TF-IDF and word-embedding as content-based filters~\cite{scikit-learn,rehurek_lrec}. Hereby, the otherwise intrinsic inductive bias is reduced. The word-embedding relies on fastText and utilizes pre-trained vectors~\cite{DBLP:journals/corr/BojanowskiGJM16}. Vector representations of items are the normalized sum of the vectors of the normalized embeddings of the individual words. In order to work with the data in an efficient way, transactions are sampled. For this study a sample-size of $100,000$ was chosen. This represents a fair balance between required computing time and ability to generalize.

				The recall in a Top-N test set is used as score of how good an algorithm is able to recommend an item.
				\begin{enumerate*}[label= (\arabic*)]
					\item For each user, first $100$ items with which the user has not interacted with are selected.
					\item The algorithm is then requested to rank an item with which the user has interacted with but which was not used for building the user's profile and the $100$ randomly sampled items. It is assumed that the random $100$ items are of low interest to the user.
					\item The position of the item with which the user has interacted with in the ranked test set may now be taken as the recall in a Top-N test set.
				\end{enumerate*}
			\end{block}
		\end{column}
	\end{columns}

@@

						\begin{figure}[ht]
							\centering
							\subcaptionbox{Interaction Frequency\label{fig:InteractionFrequency}}{
								\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/number_interactions}}}
							}
							\hspace{.5em}
							\subcaptionbox{Interaction Strength\label{fig:InteractionStrength}}{
								\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/distribtion_ratings}}}
							}
							\caption{The frequency of the number of interactions per user without outliers is shown in~\subref{fig:InteractionFrequency} and the distribution of the transformed interaction strength is shown in~\subref{fig:InteractionStrength}.}\label{fig:Interactions}
						\end{figure}
					\end{column}

					\begin{column}{0.49\textwidth}
						Statistical information about the performances of the individual machine learners in the learning subsystem are visualized in~\autoref{fig:MetaLearnerStatistics}. Overall, content-based filters are significantly better in achieving a low recall-position. The baseline is set to be the single best algorithm, i.e., TF-IDF. As seen in~\autoref{fig:MetaLearnerPerformance}, the baseline is beaten by two meta-learners.

						\begin{figure}[ht]
							\centering
							\subcaptionbox{Statistics\label{fig:MetaLearnerStatistics}}{
								\includegraphics[width=.45\textwidth,height=\textheight,keepaspectratio]{{{../res/learning_subsystem_position}}}
							}
							\hspace{.5em}
							\subcaptionbox{Perfromance\label{fig:MetaLearnerPerformance}}{
								\includegraphics[width=.45\textwidth,height=\textheight,keepaspectratio]{{{../res/meta_learner_performance_average_position}}}
							}
							\caption{Statistical features of the distributions of the recall-positions of the learning subsystem are shown in~\subref{fig:MetaLearnerStatistics}. The dashed green lines indicate the mean, the straight orange lines the median. The percentage below each algorithm indicates how often it is the best possible algorithm. To resolve a draw deterministically, the ``best'' algorithm is decided in alphabetic order.
							The performance of the meta-learner system addressing the ASP is shown in~\subref{fig:MetaLearnerPerformance}. The dashed orange line represents the overall best algorithm, i.e., TF-IDF.}\label{fig:MetaLearner}
						\end{figure}
					\end{column}
				\end{columns}
			\end{block}
		\end{column}
	\end{columns}

@@

>>>>

\usepackage[font=small,labelfont=bf]{caption}  % Customize caption aesthetics
\usepackage[subrefformat=parens,labelformat=parens,font=small]{subcaption}  % Customize sub-captions

@@

\newcommand{\mathcolorbox}[2]{\colorbox{#1}{$\displaystyle #2$}}
\newcommand{\parvspace}[1]{\vspace{.5em}}  % Manually HACK space after a paragraph

@@

% Prettify poster
\definecolor{TrinityBlue}{RGB}{32, 113, 182}
\usefonttheme[onlymath]{serif}%
\setbeamerfont{title}{size=\LARGE}
\setbeamercolor*{title}{fg = white}%
\setbeamercolor*{author}{fg = white}%
\setbeamercolor*{institute}{fg = white}%
\setbeamercolor*{date}{fg = white}%
\setbeamercolor*{title page}{fg = black}%
\setbeamercolor*{item}{fg = black}%
\setbeamercolor*{block title}{fg = TrinityBlue}%
\setbeamercolor*{caption name}{fg = black}%
\setbeamercolor*{bibliography item}{fg = black}
\setbeamercolor*{bibliography entry author}{fg = black}%
\setbeamercolor*{bibliography entry title}{fg = black}%
\setbeamercolor*{bibliography entry location}{fg = black}%
\setbeamercolor*{bibliography entry note}{fg = black}%
\setbeamerfont{title}{series = \bfseries}%
\setbeamerfont{institute}{size = \small}%
\setbeamertemplate{bibliography item}[text]
\setbeamertemplate{caption}[numbered]%
\setbeamertemplate{caption label separator}[endash]%
%\usebackgroundtemplate{\includegraphics[width=\paperwidth,height=\paperheight]{./MUW_green.pdf}}  % Background image

\title{Augmenting the DonorsChoose.org Corpus for Meta-Learning}
\subject{Algorithm Selection and Meta-Learning in Information Retrieval}
\author{Gordian~Edenhofer\inst{1,2} \and Andrew~Collins\inst{1} \and Akiko~Aizawa\inst{2} \and Joeran~Beel\inst{1,2}}
\institute{%
	\inst{1}Trinity College Dublin, School of Computer Science \& Statistics, ADAPT Centre, Ireland\\
	\inst{2}National Institute of Informatics Tokyo, Digital Content and Media Sciences Division, Japan
	\and \href{mailto:beelj@tcd.ie}{beelj@tcd.ie}, \href{mailto:collinsa@tcd.ie}{collinsa@tcd.ie}, \href{mailto:aizawa@nii.ac.jp}{aizawa@nii.ac.jp}
}
\date{\itshape{} Algorithm Selection and Meta-Learning in Information Retrieval (AMIR)}  % Set date to venue

@@

\begin{frame}[t,fragile=singleslide]{}
	% Headline
	\colorbox{TrinityBlue}{
		\hspace{-1.02em}  % Ugly HACK to center background block around header and avoid "overfull \hbox" message
		\begin{columns}[t]
			\begin{column}{0.01\textwidth}
			\end{column}

			\begin{column}{0.18\textwidth}
				\flushleft%
				\vspace{1em}
				\includegraphics[width=0.75\columnwidth]{{{../res/Logo of Trinity College Dublin}}}
			\end{column}

			\begin{column}{0.5\textwidth}
				\vspace{1em}
				\titlepage{}
			\end{column}

			\begin{column}{0.18\textwidth}
				\flushright%
				\vspace{1em}
				\vspace{2em}\includegraphics[width=0.8\columnwidth]{{{../res/Logo of ADAPT}}}
			\end{column}

			\begin{column}{0.01\textwidth}
			\end{column}
		\end{columns}
	}
	\vspace{0.em}  % HACK to leave some space between the header and the content

	\begin{columns}[t]
		\begin{column}{0.45\textwidth}
			\begin{block}{}
			The augmented dataset and the source code are released into the public domain at \colorbox{yellow}{\bfseries \href{https://github.com/BeelGroup/Augmented-DonorsChoose.org-Dataset}{GitHub:BeelGroup/Augmented-DonorsChoose.org-Dataset}}. Each feature worthy of an explanation is described in the repository.
			\end{block}
			\vspace{1.5em}

			\begin{block}{Introduction}
				Meta-Learning is the process of applying machine learning on metadata of previous machine learning experiments. The goal is to improve existing approaches by combining the strengths of several single machine learners. The metadata though needs to be computed in advance.
				\parvspace{}

				Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, the choice of the metric for computing metadata of single machine learners is important. Using the RMSE, as is done in some of the aforementioned publications, disallows using content-based methods.
				\parvspace{}

@@

		{
			\color{gray}
			\hspace{-1em}
			\vrule{}
			\hspace{+1em}
		}

		\begin{column}{0.45\textwidth}
			\begin{block}{Methodology}
				The DonorsChoose.org dataset is composed of past donations featuring $4.69$ million transactions performed by $2.12$ million users having donated to $1.11$ million projects. In contrast to, e.g., the popular MovieLense dataset, the DonorsChoose.org corpus provides significantly more transactions with more detailed item and transaction information as exemplary seen in~\autoref{tab:IllustrativeTable}.
				\parvspace{}

				The augmented data which is to the right of the separator in~\autoref{tab:IllustrativeTable} is composed of the metadata of SVD, KNN, TF-IDF and word-embedding using fastText~\cite{scikit-learn,rehurek_lrec,DBLP:journals/corr/BojanowskiGJM16}. Vector representations of items are the normalized sum of the vectors of the normalized embeddings of the individual words. In order to work with the data in an efficient way, $100,000$ transactions are sampled.
				\parvspace{}

				The recall in a Top-N test set is used as score of how good an algorithm is able to recommend an item.
				\begin{enumerate}
					\item For each user, first $100$ items with which the user has not interacted are selected.
					\item The algorithm then ranks an item with which the user has interacted and the $100$ randomly sampled items.
					\item The position of the item with which the user has interacted in the ranked test set may now be taken as the recall in a Top-N test set.
				\end{enumerate}
			\end{block}
		\end{column}
	\end{columns}

	\begin{columns}[t]
		\begin{column}{0.645\textwidth}
			\begin{table}[ht]
				\centering
				\includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{{{../res/table}}}
				\caption{Illustrative example of the overall design of the augmented transaction table with amended learning subsystem performance scores, statistics and meta-learner information.}%
				\label{tab:IllustrativeTable}
			\end{table}
		\end{column}
	\end{columns}

	\begin{columns}[t]
		\begin{column}{0.945\textwidth}
			\color{gray}
			\vspace{0.5em}
			\hrule{}
			\vspace{0.5em}
		\end{column}
	\end{columns}

@@

						\begin{figure}[ht]
							\centering
							\subcaptionbox{Interaction Frequency\label{fig:InteractionFrequency}}{
								\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/number_interactions}}}
							}
							\hspace{.5em}
							\subcaptionbox{Interaction Strength\label{fig:InteractionStrength}}{
								\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/distribtion_ratings}}}
							}
							\caption{The frequency of the number of interactions per user without outliers is shown in~\subref{fig:InteractionFrequency} and the distribution of the transformed interaction strength is shown in~\subref{fig:InteractionStrength}.}%
							\label{fig:Interactions}
						\end{figure}
					\end{column}

					\begin{column}{0.49\textwidth}
						Statistical features about the performances of the individual machine learners in the learning subsystem are visualized in~\autoref{fig:MetaLearnerStatistics}. Overall, content-based filters are significantly better in achieving a low recall-position. The baseline is set to be the single best algorithm, i.e., TF-IDF. As seen in~\autoref{fig:MetaLearnerPerformance}, the baseline is beaten by two meta-learners.

						\begin{figure}[ht]
							\centering
							\subcaptionbox{Statistics\label{fig:MetaLearnerStatistics}}{
								\includegraphics[width=.45\textwidth,height=\textheight,keepaspectratio]{{{../res/learning_subsystem_position}}}
							}
							\hspace{.5em}
							\subcaptionbox{Perfromance\label{fig:MetaLearnerPerformance}}{
								\includegraphics[width=.45\textwidth,height=\textheight,keepaspectratio]{{{../res/meta_learner_performance_average_position}}}
							}
							\caption{Statistical features of the recall-positions of the learning subsystem are shown in~\subref{fig:MetaLearnerStatistics}. The dashed green lines indicate the means, the straight orange lines the medians. The percentage below each algorithm indicates how often it is the best possible algorithm. To resolve a draw deterministically, the ``best'' algorithm is decided in alphabetic order.
							The performance of the meta-learner system is shown in~\subref{fig:MetaLearnerPerformance}. The dashed orange line represents the overall best algorithm, i.e., TF-IDF.}%
							\label{fig:MetaLearner}
						\end{figure}
					\end{column}
				\end{columns}
			\end{block}
		\end{column}
	\end{columns}

	\begin{columns}[t]
		\begin{column}{0.945\textwidth}
			\color{gray}
			\vspace{0.5em}
			\hrule{}
			\vspace{0.5em}
		\end{column}
	\end{columns}

@@

		{
			\color{gray}
			\hspace{-1em}
			\vrule{}
			\hspace{+1em}
		}
